Energy Price Forecasting.

### 2024-08-01 minos/transitions/energy_price_forecasting.Rmd
saving scatterplot for gain vs cost

```{r}
electric.series <- electricity_vector$electricity #+ rnorm(length(electric.series))
# ACF, difference 1 ACF and PACF suggests AR(2)  difference 1 and MA(4) structure (2, 1, 4) in ARIMA.
acf(electric.series)
acf(diff(electric.series))
pacf(electric.series)
#arima.electric 
#m1 <- auto.arima(electric.series)
m1 <- arima(electric.series, order=c(7, 0, 1))
print(m1)
#Box.test(residuals(m1), type="Ljung-Box")
plot(residuals(m1))
autoplot(forecast(m1, 40))
checkresiduals(m1)
print(m1$aic)
# residuals appear to be in nominal range. large LB p-value also support no autocorrelation between residuals. 
```

```{r}
gas.series <- gas_vector$gas
# ACF, difference 1 ACF and PACF suggests AR(2)  difference 1 and MA(4) structure (2, 1, 4) in ARIMA.
acf(gas.series)
acf(diff(gas.series))
pacf(gas.series)
#arima.electric 
m2 <- arima(gas.series, order=c(4, 0 ,1))
#Box.test(residuals(m1), type="Ljung-Box")
plot(residuals(m2))
autoplot(forecast(m2, 15))
checkresiduals(m2)
print(m2$aic)
# residuals appear to be in nominal range. large LB p-value also support no autocorrelation between residuals. 
```

```{r}
solid.series <- solid_fuel_vector$solid_fuel
# ACF, difference 1 ACF and PACF suggests AR(2)  difference 1 and MA(4) structure (2, 1, 4) in ARIMA.
acf(solid.series)
acf(diff(solid.series))
pacf(solid.series)
#arima.electric 
m3 <- arima(solid.series, order=c(5, 1 ,5))
#Box.test(residuals(m1), type="Ljung-Box")
plot(residuals(m3))
autoplot(forecast(m3, 15))
checkresiduals(m3)
print(m3$aic)
# residuals appear to be in nominal range. large LB p-value also support no autocorrelation between residuals. 
```

```{r}
liquid.series <- liquid_fuel_vector$liquid_fuel
# ACF, difference 1 ACF and PACF suggests AR(2)  difference 1 and MA(4) structure (5, 0, 1) in ARIMA.
acf(liquid.series)
acf(diff(liquid.series))
pacf(liquid.series)
#arima.electric 
m4 <- arima(gas.series, order=c(5, 0 ,1))
#Box.test(residuals(m1), type="Ljung-Box")
plot(residuals(m4))
autoplot(forecast(m4,15))
checkresiduals(m4)
print(m4$aic)
# residuals appear to be in nominal range. large LB p-value also support no autocorrelation between residuals. 
```

```{r}
# saving data to R objects for use in MINOS
create.if.not.exists(here::here("data", "transitions", "energy_prices")
write.csv(electric.series, "data/transitions/energy_prices/solid_historic.csv")
write.csv(gas.series, "data/transitions/energy_prices/gas_historic.csv")
write.csv(solid.series, "data/transitions/energy_prices/solid_historic.csv")
write.csv(liquid.series, "data/transitions/energy_prices/liquid_historic.csv")
write.csv(forecast(m1, 15), "data/transitions/energy_prices/electric_arima.csv")
write.csv(forecast(m2, 15), "data/transitions/energy_prices/gas_arima.csv")
write.csv(forecast(m3, 15), "data/transitions/energy_prices/solid_arima.csv")
write.csv(forecast(m4, 15), "data/transitions/energy_prices/liquid_arima.csv")
```

>>>>

```{r}
electric.series <- electricity_vector$electricity #+ rnorm(length(electric.series))
# ACF, difference 1 ACF and PACF suggests AR(2)  difference 1 and MA(4) structure (2, 1, 4) in ARIMA.
acf(electric.series)
acf(diff(electric.series))
pacf(electric.series)
#arima.electric 
#m1 <- auto.arima(electric.series)
m1 <- arima(electric.series, order=c(7, 0, 1))
print(m1)
#Box.test(residuals(m1), type="Ljung-Box")
plot(residuals(m1))
autoplot(forecast(m1, 40))
checkresiduals(m1)
print(m1$aic)

pdf(here::here("plots", "electric_forecast.pdf"))
autoplot(forecast(m1, 15))
dev.off()
# residuals appear to be in nominal range. large LB p-value also support no autocorrelation between residuals. 
```

```{r}
gas.series <- gas_vector$gas
# ACF, difference 1 ACF and PACF suggests AR(2)  difference 1 and MA(4) structure (2, 1, 4) in ARIMA.
acf(gas.series)
acf(diff(gas.series))
pacf(gas.series)
#arima.electric 
m2 <- arima(gas.series, order=c(4, 0 ,1))
#Box.test(residuals(m1), type="Ljung-Box")
plot(residuals(m2))
autoplot(forecast(m2, 15))
checkresiduals(m2)
print(m2$aic)

pdf(here::here("plots", "gas_forecast.pdf"))
autoplot(forecast(m2, 15))
dev.off()
# residuals appear to be in nominal range. large LB p-value also support no autocorrelation between residuals. 
```

```{r}
solid.series <- solid_fuel_vector$solid_fuel
# ACF, difference 1 ACF and PACF suggests AR(2)  difference 1 and MA(4) structure (2, 1, 4) in ARIMA.
acf(solid.series)
acf(diff(solid.series))
pacf(solid.series)
#arima.electric 
m3 <- arima(solid.series, order=c(5, 1 ,5))
#Box.test(residuals(m1), type="Ljung-Box")
plot(residuals(m3))
autoplot(forecast(m3, 15))
checkresiduals(m3)
print(m3$aic)
# residuals appear to be in nominal range. large LB p-value also support no autocorrelation between residuals.

pdf(here::here("plots", "solid_forecast.pdf"))
autoplot(forecast(m3, 15))
dev.off()
```

```{r}
liquid.series <- liquid_fuel_vector$liquid_fuel
# ACF, difference 1 ACF and PACF suggests AR(2)  difference 1 and MA(4) structure (5, 0, 1) in ARIMA.
acf(liquid.series)
acf(diff(liquid.series))
pacf(liquid.series)
#arima.electric 
m4 <- arima(gas.series, order=c(5, 0 ,1))
#Box.test(residuals(m1), type="Ljung-Box")
plot(residuals(m4))
autoplot(forecast(m4,15))
checkresiduals(m4)
print(m4$aic)

pdf(here::here("plots", "oil_forecast.pdf"))
autoplot(forecast(m4, 15))
dev.off()
# residuals appear to be in nominal range. large LB p-value also support no autocorrelation between residuals. 
```

```{r}
# saving data to R objects for use in MINOS
create.if.not.exists(here::here("data", "transitions", "energy_prices"))
write.csv(electric.series, here::here("data", "transitions", "energy_prices", "solid_historic.csv"))
write.csv(gas.series, here::here("data", "transitions", "energy_prices", "gas_historic.csv"))
write.csv(solid.series, here::here("data", "transitions", "energy_prices", "solid_historic.csv"))
write.csv(liquid.series, here::here("data", "transitions", "energy_prices", "liquid_historic.csv"))
write.csv(forecast(m1, 15), here::here("data", "transitions", "energy_prices", "electric_arima.csv"))
write.csv(forecast(m2, 15), here::here("data", "transitions", "energy_prices", "gas_arima.csv"))
write.csv(forecast(m3, 15), here::here("data", "transitions", "energy_prices", "solid_arima.csv"))
write.csv(forecast(m4, 15), here::here("data", "transitions", "energy_prices", "liquid_arima.csv"))
```
